### 2019-01-30 paper/response_to_referee.txt
Start of response to referee.

Re: DZ12205
   Surrogate model for an aligned-spin effective one body waveform model
   of binary neutron star inspirals using Gaussian process regression
   by Benjamin D. Lackey, Michael Pürrer, Andrea Taracchini, et al.

Dear Dr. Lackey,

The above manuscript has been reviewed by one of our referees.
Comments from the report appear below for your consideration.

When you resubmit your manuscript, please include a summary of the
changes made and a brief response to all recommendations and
criticisms.

Yours sincerely,

Joshua C. Sayre
Associate Editor
Physical Review D
Email: prd@aps.org
https://journals.aps.org/prd/

----------------------------------------------------------------------
Report of the Referee -- DZ12205/Lackey
----------------------------------------------------------------------

In this manuscript, the authors construct a surrogate for SEOBNRv4T
using Gaussian process regression (GPR). The surrogate is intended for
spin-aligned binary neutron star waveforms, but the techniques they
introduce are flexible enough to accommodate precession in the future.

One of the novel aspects of their work is to fit to residuals between
SEOBNRv4T and TaylorF2 thus reducing the requirement for very high
accuracy. They keep their TaylorF2 model as "ignorant" as SEOBNRv4T to
make the phase residual smaller as explained in App. A. The EOB model
includes dynamical tides from the induced f-mode resonances. They use
the Yagi et al. I-Love-Q relations to reduce the dimension of the
parameter space from 14 to 5 for the training sets. They then sample
this 5-D "cube" as a Latin Hyper Cube, thus an initial 128 points + 32
corners suffices to train the surrogate. The "training" over the 5
parameters is done using GPR which is the other novel contribution of
this work to the relevant literature. GPR has the additional nice
property that it "provides a convenient estimate of its own
uncertainty" to use the authors' words. The authors have the option of
retraining the surrogate using 400+400 more points in the 5-D space.
Their final choice for the best surrogate is generated from 960
training sets using a Matern kernel for the covariance. This surrogate
has a mismatch of at worst 4.5E-4 and a phase disagreement of <~ 1
radian at f <~ 2 kHz. The surrogate is finally shown to be good enough
for parameter extraction in LIGO-Virgo.

The article is well written, the figures are of high quality and the
work is original. In my opinion, this article will make a nice
addition to the existing literature on this topic. I recommend it for
publication modulo minor questions/comments/corrections that I list
below.

1) In the abstract, it is written the surrogate takes 0.13 seconds to
evaluate from 20Hz, but given the flat cost of 0.01s of GPR
interpolation at each frequency node and given the fact that there are
roughly 40 frequency according to Sec. III.C, isn't the overall run
time more like 40 x 0.01 + 0.13 = 0.53 second?

 - The confusion is due to bad phrasing on our part. When we wrote 0.01s to evaluate each node, we meant 0.01s to evaluate all 40 nodes. We have changed the wording in Section IV.B (paragraph 2) and Section VI (paragraph 2) to clarify this.

2) In the left column of page 2, the sentence following Refs. [32-36]
contains the word "instead" twice, which seems redundant.

 - We removed the second "instead".

3) In the upperleft part of page 3, it is written that "up to 0.85 for
unequal-mass (up to 0.98 for equal-mass)". Where are these values
obtained from? According to Ref[24] of this manuscript, the equal-mass
case seems to have been extended to possibly 0.994.

4) The authors may wish to add https://arxiv.org/abs/1812.02744 to
their list of papers that accommodate tidal effects to EOB in the
first paragraph of page 3.

 - We have added this reference to the second paragraph of Section II.A.

5) Immediately after Eq.(9), the authors cite Damour-Nagar-Villain
2012. It might be useful to also cite recent work of
https://arxiv.org/abs/1805.07266 that have now extended DNV.

6) Typo in the lowerleft of page 5 where the word "subsequent" has
been written as "subsequenct".

 - We have fixed this typo.

7) Could the authors comment on why they have chosen 12M instead of
15M in Eqs.(22-23)? Is the matching to omega_{match} made to be
asymptotic to ensure smoother Fourier modes?

8) Could the authors comment on the bifurcation of Phi in Fig. 1? It
seems to me that there would be such a difference between the
Lambda_A=Lambda_B=1000 corners of the parameter space vs.
Lambda_A=Lambda_B=0 corners, but the count does not quite add up. Do
these branches each contain 16 waveforms? It might be nice to add a
brief explanation about this to the caption of Fig. 1 for the sake of
the curious, but less familiar, readers.

 - The bifurcation is due to the two extremes in parameter space for q. The lower 16 curves correspond to q=1, and the upper 16 curves correspond to q=1/3. This results because the mass ratio dominates over the other 4 parameters. We added a sentence to the caption of Fig. 1 to clarify this.

9) For Sec IV. B regarding run times, could the authors explicitly
provide an overall run time for the 20Hz non-spinning BNS run
including the flat cost of ~0.01sec per frequency node for computing
the GPR fits? According to my estimation in point 1) above, this would
be 0.53 second which interestingly coincides with the value given by
Table IV of 1812.02744 in their App. B. Looking at that table begs the
question: do the authors have any idea how long their surrogate would
take to run after being trained over a set starting from 10Hz?

 - We apologize for the confusion that also resulted in comment 1 above. The times shown in Fig. 8 are in fact the total evaluation times for the waveform including evaluating the GPR fits at all nodes and resampling the final waveform at uniform frequency spacing (~0.01s for 80Hz, 0.13s for 20Hz, and ~1 s for 10Hz).

 We have clarified that the evaluation times shown are the *total* surrogate evaluation time in Section IV.B. sentence 1 of paragraph 2. In the same paragraph, we also clarified that the 0.01s is the time needed to evaluate all 40 nodes (not just 1) as discussed in comment 1 above.

10) In Sec V. the authors show that their surrogate faithfully
recovers the injected parameters. First of all, I wanted to commend
the authors for sticking with presenting the more pain-in-the-neck q=1
case despite the fact that this value sits on the upper bound of the q
prior. Have the authors repeated their parameter estimation
computations for "friendlier" values of q such as 0.8? If so, do the
PDFs look less asymmetric and more importantly are the centres of the
peaks of the PDFs closer to the injected values?

 - At the moment one of our internal LIGO reviewers is performing an analysis with a less symmetric masses. Those runs, however, are not complete.